---
title: "Summary of the statistical filters"
author: "Kyle Taysom"
date: "12/6/2020"
output: html_document
---

```{r}
require(readr)
require(ggplot2)
require(reshape2)

```

```{r}
#Load the data from outlier filtering round 1

go_1 = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/gross_outlier_summary.csv")
go_2 = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_2_data/gross_outlier_summary.csv")
go_1
```
The number of original samples
```{r}
go_1$count = go_1$Not_outliers+go_1$Outlier_count
sum(go_1$count)

go_2$count = go_2$Not_outliers+go_2$Outlier_count
sum(go_2$count)

data.frame("round_1" = sum(go_1$count), "round_2" = sum(go_2$count))
```
The number of original feed types
```{r}
data.frame("round_1" = length(unique(go_1$product)), "round_2" = length(unique(go_2$cluster)))

```
Summary of gross outliers removed
```{r}

print(summary(go_1$pct_outliers))
print(summary(go_2$pct_outliers))

```
The total % of gross outliers removed by each round
```{r}
data.frame("Round 1" = sum(go_1$Outlier_count)/sum(go_1$Not_outliers+go_1$Outlier_count), 
           "Round 2" = sum(go_2$Outlier_count)/sum(go_1$Not_outliers+go_1$Outlier_count))

```
Summarize the number of nutrients selected
- PCA nutrients were the maximum number of nutrients that could be included and still meet the minimum threshold of 30% of the samples or at least 30 samples retained
- which feed type had the most and what were they
- which feed type had the least and what were they
- what was the mean number of nutrients per feed type
- can we say something about the types of feeds that had many vs few (forages vs non) (nir vs chemistry)?
```{r}
nutrients_chosen_1 = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/PCA_feed_nutrients_chosen.csv")
nutrients_chosen_2 = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_2_data/PCA_feed_nutrients_chosen.csv")


nutrients_chosen_1$count = sapply(nutrients_chosen_1$nutrients, FUN = function(x) length(unlist(strsplit(x, split = ","))))
nutrients_chosen_2$count = sapply(nutrients_chosen_2$nutrients, FUN = function(x) length(unlist(strsplit(x, split = ","))))

nutrients_chosen_1 = nutrients_chosen_1[order(nutrients_chosen_1$count, decreasing = T),]
nutrients_chosen_2 = nutrients_chosen_2[order(nutrients_chosen_2$count, decreasing = T),]


print(paste("The feed with the most nutrients was", nutrients_chosen_1$Feed_type[1], "with", nutrients_chosen_1$count[1], "nutrients, including ",nutrients_chosen_1$nutrients[1], sep = " "))

last_feed_num = dim(nutrients_chosen_1)[1]-1
print(paste("The feed with the least nutrients was", nutrients_chosen_1$Feed_type[last_feed_num], "with", nutrients_chosen_1$count[last_feed_num], "nutrients, including ",nutrients_chosen_1$nutrients[last_feed_num], sep = " "))

print(paste("The mean number of nutrients included was ", mean(nutrients_chosen_1$count)))

#for round 2
print(paste("The feed with the most nutrients was", nutrients_chosen_2$Feed_type[1], "with", nutrients_chosen_2$count[1], "nutrients, including ",nutrients_chosen_1$nutrients[2], sep = " "))

last_feed_num = dim(nutrients_chosen_2)[1]-1
print(paste("The feed with the least nutrients was", nutrients_chosen_2$Feed_type[last_feed_num], "with", nutrients_chosen_2$count[last_feed_num], "nutrients, including ",nutrients_chosen_2$nutrients[last_feed_num], sep = " "))

print(paste("The mean number of nutrients included was ", mean(nutrients_chosen_1$count)))
```



Summarize the number of samples removed by PCA outlier in rounds 1 and 2
- Feed types were dropped when they didn't have at least 100 samples with complete cases for at least 3 nutrients
- How many feed types were dropped because they didn't have enough samples with enough nutrients?
```{r}
PCA_out_1 = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/PCA_outlier_summary.csv")
dropped_feeds_1 = unlist(PCA_out_1$product[PCA_out_1$not_PCA_outlier == 0])

PCA_out_2 = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_2_data/PCA_outlier_summary.csv")
dropped_feeds_2 = PCA_out_2$cluster[PCA_out_2$not_PCA_outlier == 0]

print(paste("Round 1 dropped",dropped_feeds_1, "for insuffiennt number of samples with complete cases", sep = " "))
print(paste("Round 2 dropped", dropped_feeds_2, "for insufficient number of samples with complete cases", sep = " "))

sum(table(round_1_all_outlier_data$product[round_1_all_outlier_data$product %in% dropped_feeds_1]))/dim(round_1_all_outlier_data[!round_1_all_outlier_data$is_a_nutrient_outlier,])[1]
```
```{r}
#Summarize the number of samples removed because they were not complete cases for PCA nutrients

round_1_all_outlier_data = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/m_after_outlier_detection.csv")
round_2_all_outlier_data = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_2_data/m_after_outlier_detection.csv")


#Of the samples that were not in feed types that were dropped
  #and were not nutrient outlieis_a_nutrient_outlier
  #how many were not complete cases

round_1_PCA_qualifying = table(round_1_all_outlier_data$is_a_complete_PCA_case[!(round_1_all_outlier_data$product %in% dropped_feeds_1) & !round_1_all_outlier_data$is_a_nutrient_outlier])

print(paste("Round 1 removed",round(round_1_PCA_qualifying[1]/sum(round_1_PCA_qualifying)*100,2),"because they did not have complete sets of the selected nutrients", sep = " "))

print(paste("In round 1",round_1_PCA_qualifying[1],"samples were dropped because they were not complete cases for the nutrients selected for PCA analysis", sep = " "))

round_2_PCA_qualifying = table(round_2_all_outlier_data$is_a_complete_PCA_case[!(round_2_all_outlier_data$cluter %in% dropped_feeds_1) & !round_2_all_outlier_data$is_a_nutrient_outlier])

print("No futher samples were dropped for being incomplete cases in Round 2")
```
What percent of samples were removed by the PCA outlier step
```{r}
print(summary(PCA_out_1$percent_PCA_outliers))
print(summary(PCA_out_2$percent_PCA_outliers))

```
The percent of samples removed as PCA outliers in each round
```{r}
data.frame("Round 1" = sum(PCA_out_1$is_PCA_outlier)/sum(PCA_out_1$not_PCA_outlier+PCA_out_1$is_PCA_outlier),
           "Round 2" = sum(PCA_out_2$is_PCA_outlier)/sum(PCA_out_2$not_PCA_outlier+PCA_out_2$is_PCA_outlier))

```
Describe the number of clusters added/removed by the clustering steps
```{r}
#Round 1
  #Created wet/dry CGF, beet pulp, and corn distillers grains
  #Created solvent/extruded SBM and canola meal
  #Removed outlier clusters from hominy, soybean hulls, roasted soybeans, corn stalks, and brewers grains

#Round 2
  #No new clusters were created or removed


#Find the number of samples after PCA outlier analysis
pre_cluster_all = dim(round_1_all_outlier_data[!round_1_all_outlier_data$is_a_nutrient_outlier &  #the total number of samples pre-clustering
                           !round_1_all_outlier_data$is_PCA_outlier,])[1]

pre_cluster_set = dim(round_1_all_outlier_data[round_1_all_outlier_data$product %in% c("Hominy","Soybean_hulls","Corn_stalks","Brewers_grain") &
                                                 !round_1_all_outlier_data$is_a_nutrient_outlier &  #the number of samples in feed types that had clusters removed
                                                 !round_1_all_outlier_data$is_PCA_outlier,])[1]


#Find the number of samples, in each feed type that had clusters removed, after cluster analysis
hominy = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/post_clustering_data/Hominy.csv")
soybean_hulls = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/post_clustering_data/Soybean_hulls.csv")
corn_stalks = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/post_clustering_data/Corn_stalks.csv")
brewers_grain = read_csv("C:/Users/Kyle/OneDrive - Dairyland Laboratories, Inc/School/Final project/round_1_data/post_clustering_data/Brewers_grain.csv")

post_cluster_count = dim(hominy)[1]+dim(soybean_hulls)[1]+dim(corn_stalks)[1]+dim(brewers_grain)[1] #the number of samples in feed types that had clusters removed - after clustering

removed_cluster_samples = pre_cluster_set - post_cluster_count

print(paste(removed_cluster_samples/pre_cluster_all, " of samples were removed during cluster analysis"))
```

